Eclipse Che for Fair and Easy Education

### 2019-05-12 che.tex
Changed KeyWords,minor spelling

\begin{abstract}
Eclipse Che is a revolutionary cloud-based IDE, inheriting the approach of building cloud native software within a 
remote environment. 
Instead of developing on a local machine, a remote development-server is accessed via a webbrowser 
and the artifacts are compiled into docker-containers.

@@

\begin{keywords}
Eclipse Che \and Cloud-Ready \and Cloud-Native \and DevOps \and Docker \and Kubernetes \and OpenShift
\end{keywords}

\section{Introduction}
\label{sec:Introduction}
Children look into the clouds and count sheep. 
Developers look into the clouds and check their applications.  
Cloud technologies are already state of the art, enabling global players like Google to run applications on a world-scale.
The market-leaders stocked up their portfolios with  a  variety  of  different  cloud  technologies,  ranging from cloud computing, cloud storage, database-as-a-service or rentable authentication. 
Modern software needs to be scalable, self-contained, and should run on any device, as well as in the cloud.  
This trend is also visible in Google searches \ref{fig:digdocker}.

@@

As cloud-technology such technologies are referred which achieve five attributes: 
On-demand self-service, broad network-access, resource pooling, rapid elasticity and measured service \cite{ME11}. 
Cloud technology usually splits into two components:  infrastructure  and  services,  where  Docker and Kubernetes are considered infrastructure and provide a  platform  to  host  services.  
Most  of  the  attributes  are (pseudo-) achievable by the infrastructure alone – such as restarting containers. 
Software running as a service can be put into two categories: 
“Cloud-Ready” – meaning that it can be run in common cloud environments, such as Kubernetes. 
The term cloud ready  does  usually  not  include  live-scalability. Cloud-readieness is mostly achieved by removing dependencies and either running on bare-linux or being ported into a docker-container. 
Lately the term “Cloud-native” came up \cite{BI17}, referring to software which is built from the very first steps to run inside the cloud. This kind of software is usually running inside a docker-container and both resilient and elastic by design \cite{TO17}.  
Resilience refers to the ability of handling failure, both from external sources and restarting/configuring themselves.  
Elasticy is the ability to request more resources on demand, performing more computations when required, but also to free the resources if not needed. 
Cloud-ready software usually achieves only a certain degree of resilience.   
Cloud-Native applications are aware of their context and collect meaningful metrics for the platform that they cannot reach other services or need more capacities \cite{MIN18}. 
The applications must also be able to use the gained resources in a productive and healthy way.  
Modern  cloud-platforms  can  perform  pseudo-scalability for cloud ready-services by starting multiple instances of the same service. 
The cloud-platform performs load-balancing between these instances. 
This naive approach can work out fine, but for example starting two services working on the same database will not yield any real scalability. 
Another common problem arises when two services are started on the same machine and need the same port, just to name two problems with cloud ready  software in a cloud environment.

As Cloud is a central element of modern IT-development, it is strange to see that it is not yet widely teached across 
universities in Germany \footnote{https://www.hochschulkompass.de/studium/studiengangsuche lists currently only 10 german universities with courses on cloud computing, about 30 courses on virtualization }.
There are courses to teach either classic virtualization or docker-basics, but the knowledge of full-stack cloudplatforms and cloud native development is sealed behind corporate doors and conference workshops.
Meanwhile,  the  annual  StackOverflow  developer-survey \cite{SO19} showed, that DevOps-engineers are not only 
among the most wanted and best paid jobs but also tend to be the happiest participants.  
Every student either wants to be happy or to be rich, making Cloud an interesting topic for their career.

@@

Eclipse Che  can be run  on any  base-technology  on  localhost – this is rather for demonstrations as well as for developers working on the Che-Code. 
When run (locally) on docker, the plugin CheDir enables a portable workspace, 
which can be used on any docker based Che.
\subsection{Technology and Workflow}
\label{subsec:Tec}
A normal IDE is software on a computer which makes programming easier than doing it in a plain text-editor and commandline. 
Common tasks for an IDE are dependency management, debugging, refactoring and auto-completion.
It is important that an IDE is not \textit{required} to develop - Vim can write Java files, the JDK can compile and the JRE can run everything manually from console. 
With that in mind,an IDE is just a toolstack to make things (a lot) more comfortable – but it also has requirements such as the JDK, the .Jars needed in the classpath etc. as they are only utilized by the IDE.
This stack can be considered a \textit{workspace} as it contains everything required to work, with the IDE making the workspace comfortable to use.
Eclipse Che does not work like a normal IDE. 
Che consists of three main components shown in Figure \ref{fig:cheworkflow}\footnote{source: \url{https://en.wikipedia.org/wiki/Eclipse_Che\#/media/File:Eclipse_Che_-_Workflow.PNG}}.

@@

The second component is the web-based IDE, which is basically a webpage performing the required REST-calls and utility-tasks for the developer. 
It is sent by the administrative server when an authorized user accesses the workspace via browser. 
These two components can be enriched with plugins.  
Common plugins are language-extensions which support syntax-highlighting etc. or package managers with their regarding lifecycles.
The difference to a normal IDE plugin is that it has two components: 
One in the workspace-API and one in the hosted web-IDE.  
Che offers a suite of common plugins dependent on the project. 
When starting a Maven-project, the workspace is initialized with Maven and Java.

The last component is the Che-administration server. 
In the administration-server the workspaces are orchestrated and monitored. 
The authorization is also done at the admin-server and not by every workspace.
The workspaces can be shipped onto multiple separate machines.
The runtimes do not necessarily need to be on the same machines as their workspace. 
Instead of the web-based IDE, a desktop IDE can be connected to the workspace. 
This is done by mounting the workspaces-filesystem in the desktop  IDE. 
This  will grant access to auto-completion and other features, but will miss Che specific helpers the web-IDE offers. 
Mounting remote file systems for development sounds lie an infinite source of problems.
This feature looks to be rather a proof of concept.

@@

The idea to not only ship code, but ship the experiment as a whole makes reproduction much easier. 
With Che (and using best practices for containers) the experiments done will be containerized and therefore reproducible. 
Even more complex environments including multiple components can be made in the same paradigms forming a pod. 
Such a pod could test applications working on an cassandra-cluster, fullstack web-applications or network-traffic-simulations.

@@

\section{Benefits for Education}
\label{sec:EvalEdu}
In addition to the previous mentioned benefits, there are more points about Che that are especially important for educational purposes, such as courses in the university, professional school or research departments.
For simplicity only universities will be addressed.
\subsection{New motivational and intuitive Courses}
Eclipse Che provides the possibility to learn primary cloud-technologies hands on and in a simple and supportive manner. 
Unfortunately, these topics are currently not widely teached at university.
Che makes it possible to teach students fundamentals of modern software development.
Depending on the courses subject, one can either provide scalable software or a running environment. 
With fitting material provided, it's easy to learn cloud native development, DevOps or Cloud-Operations. 
Even complicated topics, such as microservices and IoT-applications can easily be adaptet inside Che \cite{IV18} and performed as for example as group-projects with a common goal. 
The student will need to understand cloud-components as well as team-work to succeed with Che.

@@

This  hindrance  for  education  is  usually addressed  with  IOS-laboratories  and/or  cloud-resources paid for by university. 
The topic of socio-economic factors is beyond the scope of this work, but the impact has been analyzed by the PISA Studies \cite{PISA15}, which state that socio-economicly handicapped students fail three times more likely to achieve a basic level of performance.
According to PISA, this rate has not significantly changed since the studies started in 2006. 
Germany scored better on the impact of socioeconomic-factors on education in 2015 than it did in 2006 but is worse than the global average \cite{OECDCN15}.
The Mac-Laboratories and rented cloudspace are technically a good attempt at inclusion. 
But students which are required to get to university to do their assignments are 
less  likely  to  do  so.
Making  assignments  mandatory, providing extrinsic motivation \cite{DECI93} has usually a negative impact on the subjective experience.

>>>>

\begin{abstract}
Eclipse Che is a revolutionary cloud-based IDE, inheriting the approach of building cloud native software within a 
remote environment. 
Instead of developing on a local machine, a remote development-server is accessed via web-browser 
and the artifacts are compiled into docker-containers.

@@

\begin{keywords}
Eclipse Che \and Cloud Native \and DevOps \and Docker \and Kubernetes \and Higher Education \and Software Development
\end{keywords}

\section{Introduction}
\label{sec:Introduction}
Children look into the clouds and count sheep. 
Developers look into the clouds and check their applications.  
Cloud technologies are already state of the art, enabling global players like Google to run applications on a world-scale.
The market-leaders stocked up their portfolios with  a  variety  of  different  cloud  technologies,  ranging from cloud computing, cloud storage, database-as-a-service or rented authentication. 
Modern software needs to be scalable, self-contained, and should run on any device, as well as in the cloud.  
This trend is also visible in Google searches \ref{fig:digdocker}.

@@

As cloud-technology such technologies are referred which achieve five attributes: 
On-demand self-service, broad network-access, resource pooling, rapid elasticity and measured service \cite{ME11}. 
Cloud technology usually splits into two components:  infrastructure  and  services,  where  Docker and Kubernetes are considered infrastructure and provide a  platform  to  host  services.  
Most  of  the  attributes  are (pseudo-) achievable by the infrastructure alone – such as restarting containers. 
Software running as a service can be put into two categories: 
\textit{Cloud ready} – meaning that it can be run in common cloud environments, such as Kubernetes. 
The term cloud ready  does  usually  not  include  live-scalability. Cloud-readiness is mostly achieved by removing dependencies and either running on bare-linux or being ported into a docker-container. 
Lately the term \textit{Cloud native} came up \cite{BI17}, referring to software which is built from the very first steps to run inside the cloud. This kind of software is usually running inside a docker-container and both resilient and elastic by design \cite{TO17}.  
Resilience refers to the ability of handling failure, both from external sources and restarting/configuring themselves.  
Elasticity is the ability to request more resources on demand, performing more computations when required, but also to free the resources if not needed. 
Cloud-ready software usually achieves only a certain degree of resilience.   
Cloud-Native applications are aware of their context and collect meaningful metrics for the platform that they cannot reach other services or need more capacities \cite{MIN18}. 
The applications must also be able to use the gained resources in a productive and healthy way.  
Modern  cloud-platforms  can  perform  pseudo-scalability for cloud ready-services by starting multiple instances of the same service. 
The cloud-platform performs load-balancing between these instances. 
This naive approach can work out fine, but for example starting two services working on the same database will not yield any real scalability. 
Another common problem arises when two services are started on the same machine and need the same port, just to name two problems with cloud ready  software in a cloud environment.

As Cloud is a central element of modern IT-development, it is strange to see that it is not yet widely teached across 
universities in Germany \footnote{https://www.hochschulkompass.de/studium/studiengangsuche lists currently only 10 german universities with courses on cloud computing, about 30 courses on virtualization }.
There are courses to teach either classic virtualization or docker-basics, but the knowledge of full-stack cloud-platforms and cloud native development is sealed behind corporate doors and conference workshops.
Meanwhile,  the  annual  StackOverflow  developer-survey \cite{SO19} showed, that DevOps-engineers are not only 
among the most wanted and best paid jobs but also tend to be the happiest participants.  
Every student either wants to be happy or to be rich, making Cloud an interesting topic for their career.

@@

Eclipse Che  can be run  on any  base-technology  on  localhost – this is rather for demonstrations as well as for developers working on the Che-Code. 
When run (locally) on docker, the plugin CheDir enables a portable workspace, 
which can be used on any docker based Che.
\subsection{Technology and Workflow}
\label{subsec:Tec}
A normal IDE is software on a computer which makes programming easier than doing it in a plain text-editor and commandline. 
Common tasks for an IDE are dependency management, debugging, refactoring and auto-completion.
It is important that an IDE is not \textit{required} to develop - Vim can write Java files, the JDK can compile and the JRE can run everything manually from console. 
With that in mind,an IDE is just a tool-stack to make things (a lot) more comfortable – but it also has requirements such as the JDK, the .Jars needed in the classpath etc. as they are only utilized by the IDE.
This stack can be considered a \textit{workspace} as it contains everything required to work, with the IDE making the workspace comfortable to use.
Eclipse Che does not work like a normal IDE. 
Che consists of three main components shown in Figure \ref{fig:cheworkflow}\footnote{source: \url{https://en.wikipedia.org/wiki/Eclipse_Che\#/media/File:Eclipse_Che_-_Workflow.PNG}}.

@@

The second component is the web-based IDE, which is basically a web-page performing the required REST-calls and utility-tasks for the developer. 
It is sent by the administrative server when an authorized user accesses the workspace via browser. 
These two components can be enriched with plugins.  
Common plugins are language-extensions which support syntax-highlighting etc. or package managers with their regarding lifecycles.
The difference to a normal IDE plugin is that it has two components: 
One in the workspace-API and one in the hosted web-IDE.  
Che offers a suite of common plugins dependent on the project. 
When starting a Maven-project, the workspace is initialized with Maven and Java.

The last component is the Che-administration server. 
In the administration-server workspaces are orchestrated and monitored. 
The authorization is also done at the admin-server and not by every workspace.
The workspaces can be shipped onto multiple separate machines.
The runtimes do not necessarily need to be on the same machines as their workspace. 
Instead of the web-based IDE, a desktop IDE can be connected to the workspace. 
This is done by mounting the workspaces-filesystem in the desktop  IDE. 
This  will grant access to auto-completion and other features, but will miss Che specific helpers the web-IDE offers. 
Mounting remote file systems for development sounds lie an infinite source of problems.
This feature looks to be rather a proof of concept.

@@

The idea to not only ship code, but ship the experiment as a whole makes reproduction much easier. 
With Che (and using best practices for containers) the experiments done will be containerized and therefore reproducible. 
Even more complex environments including multiple components can be made in the same paradigms forming a pod. 
Such a pod could test applications working on an Cassandra-cluster, fullstack web-applications or network-traffic-simulations.

@@

\section{Benefits for Education}
\label{sec:EvalEdu}
In addition to the previous mentioned benefits, there are more points about Che that are especially important for educational purposes, such as courses in the university, professional school or research departments.
For simplicity only universities will be addressed.
\subsection{New motivational and intuitive Courses}
Eclipse Che provides the possibility to learn primary cloud-technologies hands on and in a simple and supportive manner. 
Unfortunately, these topics are currently not widely teached at university.
Che makes it possible to teach students fundamentals of modern software development.
Depending on the courses subject, one can either provide scalable software or a running environment. 
With fitting material provided, it's easy to learn cloud native development, DevOps or Cloud-Operations. 
Even complicated topics, such as micro-services and IoT-applications can easily be adapted inside Che \cite{IV18} and performed as for example as group-projects with a common goal. 
The student will need to understand cloud-components as well as team-work to succeed with Che.

@@

This  hindrance  for  education  is  usually addressed  with  IOS-laboratories  and/or  cloud-resources paid for by university. 
The topic of socio-economic factors is beyond the scope of this work, but the impact has been analyzed by the PISA Studies \cite{PISA15}, which state that socio-economicly handicapped students fail three times more likely to achieve a basic level of performance.
According to PISA, this rate has not significantly changed since the studies started in 2006. 
Germany scored better on the impact of socioeconomic-factors on education in 2015 than it did in 2006 but is worse than the global average \cite{OECDCN15}.
The Mac-Laboratories and rented cloud space are technically a good attempt at inclusion. 
But students which are required to get to university to do their assignments are 
less  likely  to  do  so.
Making  assignments  mandatory, providing extrinsic motivation \cite{DECI93} has usually a negative impact on the subjective experience.
